---
title: "Oyster aquaculture sentiment analysis"
author: Sofia Rodas
date: June 1, 2026
format: html
message: FALSE
warning: FALSE
---

This quarto document analyzes the differences in sentiment analysis for oyster aquaculture throughout different regions of the world. The data is downloaded through Nexis Unis with the search term "oyster aquaculture". Articles from January 1st 2024 to present articles, May 2026 articles are included in the analysis. 

### Background

The seafood industry is a prolific industry that feeds millions of people around the world. With the increasing demands on fisheries, more aquaculture operations are being established. Among these industries is the oyster aquaculture industry. Oyster aquaculture is uniquely positioned as a potential to both mitigate against climate change by providing coastline protection and as a means of feeding our growing populations. 

However, some potential drawbacks to oyster aquaculture operations exist. Oyster aquaculture operations can disturb the natural ecosystems if the operation includes heavy equipment. Additionally, oysters reared in aquaculture could be more susceptible to disease than natural oysters. This counters the potential for these oysters to increase seafood production.

My interest is in analyzing whether specific regions have more negative or more positive sentiments about oyster aquaculture. 

### Nexis Unis

Nexis Unis oyster aquaculture data was downloaded for the US, Europe, Eastern Asia, and Latin America. These four regions were analyzed to assess the sentiments of oyster aquaculture and compare findings by region. These regions were selected for this analysis because the top five countries for oyster aquaculture production are China, South Korea, Japan, US, and France. Hence these are areas of interest in the topic of oyster aquaculture and the regions for which text was procured.  

One limitation of this study is that all the articles must be in English to properly analyze them. There are great discrepancies in the number of articles for each region. These findings are nevertheless interesting as they show what types of texts are translated. In other words, the translated text acts as an analysis of what the perception of the sentiment towards oyster aquaculture is for these regions from a US scholar perception and not necessarily the true sentiment these regions possess. 

#### Analysis scale

Articles analyzed from the United States: 422 articles

Articles analyzed from Europe: 134 articles

Articles analyzed from Eastern Asia: 80 articles



```{r}
# load necessary libraries
library(atrrr)
library(LexisNexisTools)
library(tidyverse)
library(quanteda)
library(tm)
library(topicmodels)
library(topicdoc)   
library(tidyverse)
library(tidytext)
library(reshape2)
library(tictoc)
library(LDAvis)
library(tsne)
library(servr)
library(sentimentr)
```

# Analyze United States text sentiment for oyster aquaculture

Since we have the most data from the United States we'll complete the text analysis on this data and choose the best lexicon-based approach. This approach will then be applied to the articles from the other regions. Using the same lexicon-based approach is important to ensure mis-characterizations are not occuring simply by choosing a different lexicon model for different regions. 

### Read in the data
```{r}
# extract data from the files
pre_files <- list.files(
  path = here::here("data", "oysteraqua_us"),   
  pattern = "\\.docx$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

pre_dat <- lnt_read(pre_files)

pre_meta_df       <- pre_dat@meta
pre_articles_df   <- pre_dat@articles
pre_paragraphs_df <- pre_dat@paragraphs

df <- tibble(
  date     = pre_meta_df$Date,
  headline = pre_meta_df$Headline,
  id       = pre_meta_df$ID,
  text     = pre_articles_df$Article,
  source   = pre_meta_df$Newspaper
)
```

## Build and clean our corpus

First we create a corpus from our articles and clean the data. The primary intention is to make the text usable for the sentiment analysis. We also remove stop words that are specific to this topic to make the analysis more complete and accurate. 

```{r corpus}
us_corpus <- corpus(x=df, text_field = "text")
```

## Tokenize and Clean

I removed common journalistic words as well as words that first appeared in my categories that were unhelpful in understanding the topics discussed. 

```{r tokens}
add_stops <- c(quanteda::stopwords("en", source = "smart"),
               "said", "says", "told", "according", 
               "`", "1", "vol", "pg", "fig", "2", "Fig", "Vol")

toks <- tokens(us_corpus, remove_punct = TRUE) |> 
  tokens_tolower()

toks1 <- tokens_select(toks, pattern = add_stops, selection = "remove")
```

### Bing lexicon sentiments
```{r}
bing <- get_sentiments("bing")

#score corpus with Bing
bing_scores <- us_corpus |> 
  tidy() |>
  unnest_tokens(word, text) |> 
  inner_join(bing, by = "word") |> 
  count(id, sentiment) |> 
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |> 
  mutate(score = positive - negative)

# join scores to corpus
us_corpus <- us_corpus |> 
  tidy() |>
  left_join(bing_scores |> 
              select(id, score), by = "id") |> 
  mutate(bing_score = replace_na(score, 0)) 


```

Visualize the positive and negative sentiments based on the bing scores.

```{r}
# visualize bing sentiment about oyster aquaculture
bing_scores |> 
  summarise(positive = sum(positive), negative = sum(negative)) |>
  pivot_longer(everything(), names_to = "sentiment", values_to = "n") |>
  ggplot() + 
  geom_col(aes(reorder(sentiment, n), n),
           fill = "#161B33") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "United States Bing Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```

### AFINN lexicon
```{r}
afinn <- get_sentiments("afinn")

#score text
afinn_scores <- us_corpus |> 
  unnest_tokens(word, text) |> 
  inner_join(afinn, by = "word") |> 
  group_by(id)|> 
  summarise(affin_score = sum(value))


#join scores to corpus
us_corpus <- us_corpus |>
  left_join(afinn_scores, by ="id") 

# remove na's from the afin scores column
us_corpus <- us_corpus |> 
  mutate(affin_score = replace_na(affin_score, 0))
```

Visualize AFINN score 
```{r}
# assign positive and negative to adinn scores based on the value
us_corpus$affin_sent <-  if_else(us_corpus$affin_score > 0, "Positive", "Negative")

# visualize sentiment about oyster aquaculture
us_corpus |> 
  count(affin_sent) |> 
ggplot() + 
  geom_col(aes(affin_sent, n),
           fill = "#161B33") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "United States AFIN Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```


### NRC lexicon
```{r}
nrc <- get_sentiments("nrc")

# calculate the nrc scores
nrc_scores <- us_corpus |> 
  unnest_tokens(word, text) |> 
  inner_join(nrc, by = "word") |> 
  group_by(id) |> 
  count(sentiment)

#join scores to post
us_corpus <- us_corpus |> 
  left_join(nrc_scores, by = "id")

# replace na scores with zero
us_corpus <- us_corpus |> 
  mutate(nrc_score = replace_na(n, 0)) |> 
  select(-n)
```

Visualize the sentiments found in the text discussing oyster aquaculture in the United States.
```{r}
#visualize bluesky sentiment about sea level rise
ggplot(nrc_scores) + 
  geom_col(aes(reorder(sentiment, n), n),
           fill = "#161B33") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "United States NRC Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```
## VADER score
```{r}
us_corpus <- us_corpus |> 
  mutate(vader_score = purrr::map(text, ~ tidyvader::vader_chr(.x)$compound))
```

Visualize the VADER score.
```{r}
# assign positive and negative to adinn scores based on the value
us_corpus$vader_sent <-  if_else(us_corpus$vader_score > 0, "Positive", "Negative")


# visualize sentiment about oyster aquaculture
us_corpus |> 
  count(vader_sent) |> 
ggplot() + 
  geom_col(aes(vader_sent, n),
           fill = "#161B33") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "United States VADER Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```

# Explore classification differences between the different lexicon sentiment analysis

For this analysis, we start by analyzing the differences between the BING and the AFINN scores. When the scores do not match, one score being negative and one score being positive, let's analyze which is more accurate. We can also analyze what the NRC and VADER score show. Then we'll pick the best model and apply it to the text for Eastern Asian and Europe.

```{r}
# select those texts where the bing and afinn score don't match
diff_bing_affin <- us_corpus |> 
  mutate(bing_categ = ifelse(bing_score > 0, "positive", "negative"),
affin_categ = ifelse(affin_score > 0, "positive", "negative")) |> 
  filter(bing_categ != affin_categ)

# calculate absolute difference in bing and afinn scores
diff_reorder<- diff_bing_affin |> 
  mutate(diff_bing_affin = abs(bing_score - affin_score)) |> 
  arrange(desc(diff_bing_affin))

# calculate percent of positive nrc scores
nrc_posneg <- diff_reorder |> 
  filter(sentiment == "positive" | sentiment == "negative") |> 
  group_by(id, sentiment) |> 
  summarize(n = sum(nrc_score)) |> 
  ungroup() |> 
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |>
  mutate(nrc_positive_pct = positive / (positive + negative))

# add the percent positive nrc scores to the diff_reorder df
diff_reorder <- diff_reorder |> 
  left_join(nrc_posneg, by = "id")

```

### Explore the greatest mis-classifications and determine which is classification is best. 

```{r}
print(substr(diff_reorder$text[1], start = 1, stop = 500))
paste0("The BING score is ", diff_reorder$bing_score[1])
paste0("The AFINN score is ", diff_reorder$affin_score[1])
paste0("The NRC score is ", diff_reorder$nrc_positive_pct[1])
paste0("The VADER score is ", diff_reorder$vader_score[1])

```

Overall the text is scientific. The BING score is negative while all of the other scores are positive. Both reading the text and the majority of the sentiment scores being negative indicate that the BING score is wrong. For this specific example I would choose the NRC score. The NRC score shows the text is a little over half positive. This seems appropriate for the text. I am biased to choosing the NRC score when it performs similarly to the other lexicon sentiment scores because it allows us to investigate further sentiments with it's eight category classifications. 


```{r}
print(substr(diff_reorder$text[11], start = 1, stop = 50000))
paste0("The BING score is ", diff_reorder$bing_score[11])
paste0("The AFINN score is ", diff_reorder$affin_score[11])
paste0("The NRC score is ", diff_reorder$nrc_positive_pct[11])
paste0("The VADER score is ", diff_reorder$vader_score[11])

```


```{r}
print(substr(diff_reorder$text[21], start = 1, stop = 50000))
paste0("The BING score is ", diff_reorder$bing_score[21])
paste0("The AFINN score is ", diff_reorder$affin_score[21])
paste0("The NRC score is ", diff_reorder$nrc_positive_pct[21])
paste0("The VADER score is ", diff_reorder$vader_score[21])

```


```{r}
print(substr(diff_reorder$text[41], start = 1, stop = 50000))
paste0("The BING score is ", diff_reorder$bing_score[41])
paste0("The AFINN score is ", diff_reorder$affin_score[41])
paste0("The NRC score is ", diff_reorder$nrc_positive_pct[41])
paste0("The VADER score is ", diff_reorder$vader_score[41])
```

```{r}
print(substr(diff_reorder$text[51], start = 1, stop = 50000))
paste0("The BING score is ", diff_reorder$bing_score[51])
paste0("The AFINN score is ", diff_reorder$affin_score[51])
paste0("The NRC score is ", diff_reorder$nrc_positive_pct[51])
paste0("The VADER score is ", diff_reorder$vader_score[51])
```
# Choosing the best lexicon for the analysis of oyster aquaculture text

No lexicon analysis tool is perfect. We select the lexicon analysis approach that best characterizes the text. Based on the analysis above BING is the best generalizable lexicon analysis tool. We'll compare the BING scores for the different regions.

# Europe text sentiment analysis about oyster aquaculture

### Read in the data
```{r}
# extract data from the files
pre_files <- list.files(
  path = here::here("data", "oysteraqua_europe"),   
  pattern = "\\.docx$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

pre_dat <- lnt_read(pre_files)

pre_meta_df       <- pre_dat@meta
pre_articles_df   <- pre_dat@articles
pre_paragraphs_df <- pre_dat@paragraphs

df <- tibble(
  date     = pre_meta_df$Date,
  headline = pre_meta_df$Headline,
  id       = pre_meta_df$ID,
  text     = pre_articles_df$Article,
  source   = pre_meta_df$Newspaper
)
```

## Build and clean our corpus

First we create a corpus from our articles and clean the data. The primary intention is to make the text usable for the sentiment analysis. We also remove stop words that are specific to this topic to make the analysis more complete and accurate. 

```{r corpus}
euro_corpus <- corpus(x=df, text_field = "text")
```

## Tokenize and Clean

I removed common journalistic words as well as words that first appeared in my categories that were unhelpful in understanding the topics discussed. 

```{r tokens}
add_stops <- c(quanteda::stopwords("en", source = "smart"),
               "said", "says", "told", "according", 
               "`", "1", "vol", "pg", "fig", "2", "Fig", "Vol")

toks <- tokens(euro_corpus, remove_punct = TRUE) |> 
  tokens_tolower()

toks1 <- tokens_select(toks, pattern = add_stops, selection = "remove")
```

### Bing lexicon sentiments
```{r}
bing <- get_sentiments("bing")

#score corpus with Bing
bing_scores <- euro_corpus |> 
  tidy() |>
  unnest_tokens(word, text) |> 
  inner_join(bing, by = "word") |> 
  count(id, sentiment) |> 
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |> 
  mutate(score = positive - negative)

# join scores to corpus
euro_corpus <- euro_corpus |> 
  tidy() |>
  left_join(bing_scores |> 
              select(id, score), by = "id") |> 
  mutate(bing_score = replace_na(score, 0)) 


```

Visualize the positive and negative sentiments based on the bing scores.

```{r}
# visualize bing sentiment about oyster aquaculture
bing_scores |> 
  summarise(positive = sum(positive), negative = sum(negative)) |>
  pivot_longer(everything(), names_to = "sentiment", values_to = "n") |>
  ggplot() + 
  geom_col(aes(reorder(sentiment, n), n),
           fill = "#E8B4BC") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "Europe BING Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```


### NRC lexicon
```{r}
nrc <- get_sentiments("nrc")

# calculate the nrc scores
nrc_scores <- euro_corpus |> 
  unnest_tokens(word, text) |> 
  inner_join(nrc, by = "word") |> 
  group_by(id) |> 
  count(sentiment)

#join scores to post
euro_corpus <- euro_corpus |> 
  left_join(nrc_scores, by = "id")

# replace na scores with zero
euro_corpus <- euro_corpus |> 
  mutate(nrc_score = replace_na(n, 0)) |> 
  select(-n)
```
Visualize the sentiments found in the text discussing oyster aquaculture in Europe.
```{r}
#visualize bluesky sentiment about sea level rise
ggplot(nrc_scores) + 
  geom_col(aes(reorder(sentiment, n), n),
           fill = "#E8B4BC") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "Europe NRC Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```

# Eastern Asia text sentiment analysis about oyster aquaculture

### Read in the data
```{r}
# extract data from the files
pre_files <- list.files(
  path = here::here("data", "oysteraqua_easternasia"),   
  pattern = "\\.docx$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

pre_dat <- lnt_read(pre_files)

pre_meta_df       <- pre_dat@meta
pre_articles_df   <- pre_dat@articles
pre_paragraphs_df <- pre_dat@paragraphs

df <- tibble(
  date     = pre_meta_df$Date,
  headline = pre_meta_df$Headline,
  id       = pre_meta_df$ID,
  text     = pre_articles_df$Article,
  source   = pre_meta_df$Newspaper
)
```

## Build and clean our corpus

First we create a corpus from our articles and clean the data. The primary intention is to make the text usable for the sentiment analysis. We also remove stop words that are specific to this topic to make the analysis more complete and accurate. 

```{r corpus}
eas_corpus <- corpus(x=df, text_field = "text")
```

## Tokenize and Clean

I removed common journalistic words as well as words that first appeared in my categories that were unhelpful in understanding the topics discussed. 

```{r tokens}
add_stops <- c(quanteda::stopwords("en", source = "smart"),
               "said", "says", "told", "according", 
               "`", "1", "vol", "pg", "fig", "2", "Fig", "Vol")

toks <- tokens(eas_corpus, remove_punct = TRUE) |> 
  tokens_tolower()

toks1 <- tokens_select(toks, pattern = add_stops, selection = "remove")
```

### Bing lexicon sentiments
```{r}
bing <- get_sentiments("bing")

#score corpus with Bing
bing_scores <- eas_corpus |> 
  tidy() |>
  unnest_tokens(word, text) |> 
  inner_join(bing, by = "word") |> 
  count(id, sentiment) |> 
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |> 
  mutate(score = positive - negative)

# join scores to corpus
eas_corpus <- eas_corpus |> 
  tidy() |>
  left_join(bing_scores |> 
              select(id, score), by = "id") |> 
  mutate(bing_score = replace_na(score, 0)) 


```

Visualize the positive and negative sentiments based on the bing scores.

```{r}
# visualize bing sentiment about oyster aquaculture
bing_scores |> 
  summarise(positive = sum(positive), negative = sum(negative)) |>
  pivot_longer(everything(), names_to = "sentiment", values_to = "n") |>
  ggplot() + 
  geom_col(aes(reorder(sentiment, n), n),
           fill = "#5A6650") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "East Asia BING Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```
### NRC lexicon
```{r}
nrc <- get_sentiments("nrc")

# calculate the nrc scores
nrc_scores <- eas_corpus |> 
  unnest_tokens(word, text) |> 
  inner_join(nrc, by = "word") |> 
  group_by(id) |> 
  count(sentiment)

#join scores to post
eas_corpus <- eas_corpus |> 
  left_join(nrc_scores, by = "id")

# replace na scores with zero
eas_corpus <- eas_corpus |> 
  mutate(nrc_score = replace_na(n, 0)) |> 
  select(-n)
```
Visualize the sentiments found in the text discussing oyster aquaculture in Europe.
```{r}
#visualize bluesky sentiment about sea level rise
ggplot(nrc_scores) + 
  geom_col(aes(reorder(sentiment, n), n),
           fill = "#5A6650") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "East Asia NRC Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```

# References

Ólafsson, B. (2022). What are the environmental impacts of oysters? Retrieved from https://sentientmedia.org/oyster-farming/  




