---
title: "Oyster aquaculture sentiment analysis"
author: Sofia Rodas
date: June 1, 2026
format: html
message: FALSE
warning: FALSE
---

This quarto document analyzes the differences in sentiment analysis for oyster aquaculture throughout different regions of the world. The data is downloaded through Nexis Unis with the search term "oyster aquaculture". Articles from January 1st 2024 to present articles, May 2026 articles are included in the analysis. 

### Background

The seafood industry is a prolific industry that feeds millions of people around the world. With the increasing demands on fisheries, more aquaculture operations are being established. Among these industries is the oyster aquaculture industry. Oyster aquaculture is uniquely positioned as a potential to both mitigate against climate change by providing coastline protection and as a means of feeding our growing populations. 

However, some potential drawbacks to oyster aquaculture operations exist. Oyster aquaculture operations can disturb the natural ecosystems if the operation includes heavy equipment. Additionally, oysters reared in aquaculture could be more susceptible to disease than natural oysters. This counters the potential for these oysters to increase seafood production.

My interest is in analyzing whether specific regions have more negative or more positive sentiments about oyster aquaculture. 

### Nexis Unis

Nexis Unis oyster aquaculture data was downloaded for the US, Europe, Eastern Asia, and Latin America. These four regions were analyzed to assess the sentiments of oyster aquaculture and compare findings by region. These regions were selected for this analysis because the top five countries for oyster aquaculture production are China, South Korea, Japan, US, and France. Hence these are areas of interest in the topic of oyster aquaculture. 

```{r}
# load necessary libraries
library(atrrr)
library(LexisNexisTools)
library(tidyverse)
library(quanteda)
library(tm)
library(topicmodels)
library(topicdoc)   
library(tidyverse)
library(tidytext)
library(reshape2)
library(tictoc)
library(LDAvis)
library(tsne)
library(servr)
library(sentimentr)
```


```{r}
# extract data from the files
pre_files <- list.files(
  path = here::here("data", "oysteraqua_easternasia"),   
  pattern = "\\.docx$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

pre_dat <- lnt_read(pre_files)

pre_meta_df       <- pre_dat@meta
pre_articles_df   <- pre_dat@articles
pre_paragraphs_df <- pre_dat@paragraphs

df <- tibble(
  date     = pre_meta_df$Date,
  headline = pre_meta_df$Headline,
  id       = pre_meta_df$ID,
  text     = pre_articles_df$Article,
  source   = pre_meta_df$Newspaper
)
```

**1. Build and clean your corpus**
Create a corpus from your articles. Clean appropriately for your data —
consider what stopwords, minimum document frequencies, and tokenization
choices make sense given your specific corpus.

## Build the Corpus

We use `corpus()` from {quanteda} to build the corpus from our data
frame.

```{r corpus}
ea_corpus <- corpus(x=df, text_field = "text")
```

## Tokenize and Clean

I removed common journalistic words as well as words that first appeared in my categories that were unhelpful in understanding the topics discussed. 

```{r tokens}
add_stops <- c(quanteda::stopwords("en", source = "smart"),
               "said", "says", "told", "according", 
               "`", "1", "vol", "pg", "fig", "2")

toks <- tokens(ea_corpus, remove_punct = TRUE) |> 
  tokens_tolower()

toks1 <- tokens_select(toks, pattern = add_stops, selection = "remove")
```

I don't think i need to do this. I think the sentiment analysis works on the words not the dfm. 
Build the Document-Feature Matrix

I built a document-feature matrix (DFM) to remove words that are used in 2 or fewer articles.

```{r dfm}
# dfm1 <- dfm(toks1)
# dfm2 <- dfm_trim(dfm1, min_docfreq = 2)
# 
# # Remove empty rows — LDA will error on these
# sel_idx <- slam::row_sums(dfm2) > 0
# 
# dfm <- dfm2[sel_idx,]

```


### Bing lexicon sentiments
```{r}
bing <- get_sentiments("bing")

#score posts with Bing
bing_scores <- ea_corpus |> 
  tidy() |>
  unnest_tokens(word, text) |> 
  inner_join(bing, by = "word") |> 
  count(id, sentiment) |> 
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |> 
  mutate(score = positive - negative)

# join scores to posts
ea_corpus_scores <- ea_corpus |> 
  tidy() |>
  left_join(bing_scores |> 
              select(id, score), by = "id") |> 
  mutate(bing_score = replace_na(score, 0)) 


```


```{r nrc-plot}
# visualize sentiment about oyster aquaculture
bing_scores |> 
  summarise(positive = sum(positive), negative = sum(negative)) |>
  pivot_longer(everything(), names_to = "sentiment", values_to = "n") |>
  ggplot() + 
  geom_col(aes(reorder(sentiment, n), n),
           fill = "darkblue") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "East Asian Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```

### AFINN lexicon
```{r}
afinn <- get_sentiments("afinn")

#score text
afinn_scores <- ea_corpus |> 
  tidy() |> 
  unnest_tokens(word, text) |> 
  inner_join(afinn, by = "word") |> 
  group_by(id)|> 
  summarise(affin_score = sum(value))


#join scores to posts  
ea_corpus_scores <- ea_corpus |>
  tidy() |> 
  left_join(afinn_scores, by ="id") 

# remove na's from the afin scores column
ea_corpus_scores <- ea_corpus_scores |> 
  mutate(affin_score = replace_na(affin_score, 0))

```

### NRC lexicon
```{r}
nrc <- get_sentiments("nrc")

# calculate the nrc scores
nrc_scores <- ea_corpus |> 
  tidy() |> 
  unnest_tokens(word, text) |> 
  inner_join(nrc, by = "word") |> 
  group_by(id) |> 
  count(sentiment)

#join scores to post
ea_corpus_scores <- ea_corpus |> 
  tidy() |> 
  left_join(nrc_scores, by = "id")

# replace na scores with zero
ea_corpus_scores <- ea_corpus_scores |> 
  mutate(nrc_score = replace_na(n, 0)) |> 
  select(-n)
```
```{r}
#visualize bluesky sentiment about sea level rise
ggplot(nrc_scores) + 
  geom_col(aes(reorder(sentiment, n), n),
           fill = "darkblue") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "East Asia Sentiment about 'Oyster Aquaculture",
       y = "",
       x = "Sentiment")
```

