---
title: "Oyster aquaculture sentiment analysis"
author: Sofia Rodas
date: June 1, 2026
format: html
message: FALSE
warning: FALSE
---

This quarto document analyzes the differences in sentiment analysis for oyster aquaculture throughout different regions of the world. The data is downloaded through Nexis Unis with the search term "oyster aquaculture". Articles from January 1st 2024 to present articles, May 2026 articles are included in the analysis. 

### Background

The seafood industry is a prolific industry that feeds millions of people around the world. With the increasing demands on fisheries, more aquaculture operations are being established. Among these industries is the oyster aquaculture industry. Oyster aquaculture is uniquely positioned as a potential to both mitigate against climate change by providing coastline protection and as a means of feeding our growing populations. 

However, some potential drawbacks to oyster aquaculture operations exist. Oyster aquaculture operations can disturb the natural ecosystems if the operation includes heavy equipment. Additionally, oysters reared in aquaculture could be more susceptible to disease than natural oysters. This counters the potential for these oysters to increase seafood production.

My interest is in analyzing whether specific regions have more negative or more positive sentiments about oyster aquaculture. 

### Nexis Unis

Nexis Unis oyster aquaculture data was downloaded for the US, Europe, Eastern Asia, and Latin America. These four regions were analyzed to assess the sentiments of oyster aquaculture and compare findings by region. These regions were selected for this analysis because the top five countries for oyster aquaculture production are China, South Korea, Japan, US, and France. Hence these are areas of interest in the topic of oyster aquaculture and the regions for which text was procured.  

One limitation of this study is that all the articles must be in English to properly analyze them. There are great discrepancies in the number of articles for each region. These findings are nevertheless interesting as they show what types of texts are translated. In other words, the translated text acts as an analysis of what the perception of the sentiment for these regions is from a US scholar perception and not necessarily the true sentiment itself. 

```{r}
# load necessary libraries
library(atrrr)
library(LexisNexisTools)
library(tidyverse)
library(quanteda)
library(tm)
library(topicmodels)
library(topicdoc)   
library(tidyverse)
library(tidytext)
library(reshape2)
library(tictoc)
library(LDAvis)
library(tsne)
library(servr)
library(sentimentr)
```

# Analyze United States text sentiment for oyster aquaculture
```{r}
# extract data from the files
pre_files <- list.files(
  path = here::here("data", "oysteraqua_us"),   
  pattern = "\\.docx$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

pre_dat <- lnt_read(pre_files)

pre_meta_df       <- pre_dat@meta
pre_articles_df   <- pre_dat@articles
pre_paragraphs_df <- pre_dat@paragraphs

df <- tibble(
  date     = pre_meta_df$Date,
  headline = pre_meta_df$Headline,
  id       = pre_meta_df$ID,
  text     = pre_articles_df$Article,
  source   = pre_meta_df$Newspaper
)
```

**1. Build and clean your corpus**
First we create a corpus from our articles and clean the data. The primary 

## Build the Corpus

We use `corpus()` from {quanteda} to build the corpus from our data
frame.

```{r corpus}
us_corpus <- corpus(x=df, text_field = "text")
```

## Tokenize and Clean

I removed common journalistic words as well as words that first appeared in my categories that were unhelpful in understanding the topics discussed. 

```{r tokens}
add_stops <- c(quanteda::stopwords("en", source = "smart"),
               "said", "says", "told", "according", 
               "`", "1", "vol", "pg", "fig", "2", "Fig")

toks <- tokens(us_corpus, remove_punct = TRUE) |> 
  tokens_tolower()

toks1 <- tokens_select(toks, pattern = add_stops, selection = "remove")
```

I don't think i need to do this. I think the sentiment analysis works on the words not the dfm. 
Build the Document-Feature Matrix

I built a document-feature matrix (DFM) to remove words that are used in 2 or fewer articles.

```{r dfm}
# dfm1 <- dfm(toks1)
# dfm2 <- dfm_trim(dfm1, min_docfreq = 2)
# 
# # Remove empty rows — LDA will error on these
# sel_idx <- slam::row_sums(dfm2) > 0
# 
# dfm <- dfm2[sel_idx,]

```


### Bing lexicon sentiments
```{r}
bing <- get_sentiments("bing")

#score corpus with Bing
bing_scores <- us_corpus |> 
  tidy() |>
  unnest_tokens(word, text) |> 
  inner_join(bing, by = "word") |> 
  count(id, sentiment) |> 
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |> 
  mutate(score = positive - negative)

# join scores to corpus
us_corpus <- us_corpus |> 
  tidy() |>
  left_join(bing_scores |> 
              select(id, score), by = "id") |> 
  mutate(bing_score = replace_na(score, 0)) 


```

Visualize the positive and negative sentiments based on the bing scores.

```{r}
# visualize bing sentiment about oyster aquaculture
bing_scores |> 
  summarise(positive = sum(positive), negative = sum(negative)) |>
  pivot_longer(everything(), names_to = "sentiment", values_to = "n") |>
  ggplot() + 
  geom_col(aes(reorder(sentiment, n), n),
           fill = "darkblue") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "United States Bing Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```

### AFINN lexicon
```{r}
afinn <- get_sentiments("afinn")

#score text
afinn_scores <- us_corpus |> 
  unnest_tokens(word, text) |> 
  inner_join(afinn, by = "word") |> 
  group_by(id)|> 
  summarise(affin_score = sum(value))


#join scores to corpus
us_corpus <- us_corpus |>
  left_join(afinn_scores, by ="id") 

# remove na's from the afin scores column
us_corpus <- us_corpus |> 
  mutate(affin_score = replace_na(affin_score, 0))
```

Visualize AFINN score 
```{r}
# assign positive and negative to adinn scores based on the value
us_corpus$affin_sent <-  if_else(us_corpus$affin_score > 0, "Positive", "Negative")

# visualize sentiment about oyster aquaculture
us_corpus |> 
  count(affin_sent) |> 
ggplot() + 
  geom_col(aes(affin_sent, n),
           fill = "darkblue") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "United States AFIN Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```


### NRC lexicon
```{r}
nrc <- get_sentiments("nrc")

# calculate the nrc scores
nrc_scores <- us_corpus |> 
  unnest_tokens(word, text) |> 
  inner_join(nrc, by = "word") |> 
  group_by(id) |> 
  count(sentiment)

#join scores to post
us_corpus <- us_corpus |> 
  left_join(nrc_scores, by = "id")

# replace na scores with zero
us_corpus <- us_corpus |> 
  mutate(nrc_score = replace_na(n, 0)) |> 
  select(-n)
```

Visualize the sentiments found in the text discussing oyster aquaculture in the United States.
```{r}
#visualize bluesky sentiment about sea level rise
ggplot(nrc_scores) + 
  geom_col(aes(reorder(sentiment, n), n),
           fill = "darkblue") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "United States NRC Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```
## VADER score
```{r}
us_corpus <- us_corpus |> 
  mutate(vader_score = purrr::map(text, ~ tidyvader::vader_chr(.x)$compound))
```

Visualize the VADER score.
```{r}
# assign positive and negative to adinn scores based on the value
us_corpus$vader_sent <-  if_else(us_corpus$vader_score > 0, "Positive", "Negative")


# visualize sentiment about oyster aquaculture
us_corpus |> 
  count(vader_sent) |> 
ggplot() + 
  geom_col(aes(vader_sent, n),
           fill = "darkblue") +
  coord_flip() + 
  theme_minimal() + 
  labs(title = "United States VADER Sentiment about 'Oyster Aquaculture'",
       y = "",
       x = "Sentiment")
```

# Explore classification differences between the different lexicon sentiment analysis

For this analysis, we start by analyzing the differences between the BING and the AFINN scores. When the scores do not match, one score being negative and one score being positive, let's analyze which is more accurate. We can also analyze what the NRC and VADER score show. 

```{r}
# select those texts where the bing and afinn score don't match
diff_bing_affin <- us_corpus |> 
  mutate(bing_categ = ifelse(bing_score > 0, "positive", "negative"),
affin_categ = ifelse(affin_score > 0, "positive", "negative")) |> 
  filter(bing_categ != affin_categ)

# calculate absolute difference in bing and afinn scores
diff_reorder<- diff_bing_affin |> 
  mutate(diff_bing_affin = abs(bing_score - affin_score)) |> 
  arrange(desc(diff_bing_affin))

# calculate average NRC scores
diff_reorder <- diff_reorder |> 
                          group_by(id) |> 
                          mutate(nrc_avg = mean(nrc_score))
```

Explore the greatest mis-classifications and determine which is classification is best
```{r}
diff_reorder$text[1]
paste0("The BING score is ", diff_reorder$bing_score[1])
paste0("The AFINN score is ", diff_reorder$affin_score[1])
paste0("The NRC score is ", diff_reorder$nrc_avg[1])
paste0("The VADER score is ", diff_reorder$vader_score[1])

```

Overall the text is scientific. 
